Topic modeling on EU and the USA research projects

### 2017-06-23 thesis_report/draft_thesis.tex
EU evolution of topics over time & added code in the appendix

\setlength{\parskip}{1em}
\usepackage[justification=centering]{caption}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{cite}
\usepackage{algorithm2e}
\usepackage{url}
\usepackage{subfig}

@@

\begin{enumerate}
\item Pick a document \textbf{w} with probability $p($\textbf{w}$)$
\item For each of the $N$ words in \textbf{w}:
\begin{description}
\item (a) Pick a topic $z$ with probability $p(z|$\textbf{w}$)$
\end{description}
\begin{description}
\item (b) Generate a word $w$ with probability $p(w|z)$
\end{description}
\end{enumerate}
\begin{center}
\begin{figure}[h]
\centering
\includegraphics[width=0.6\textwidth, natwidth=548,natheight=201]{figs/PLSI_plate_notation.png}
\caption{Plate notation representing the PLSI model. \textbf{w} is the document 
index, $z$ is the topic drawn for the word $w$ from the topic distribution of 
the document}
\end{figure}
\end{center}
The result from PLSI is that every document is represented as mixing proportions
for the topics, given by $p(z|$\textbf{w}$)$. Even though PLSI is generative for 
the analyzed corpus, it is not generative for new documents, which means is that
there is no clear way of assigning probability to a document that is not part
of the training data.\cite{blei2003latent} Another problem is that the number 
of parameters in the model grows linearly with the number of documents.

@@

A limitation with the standard LDA model is the fix vocabulary. This means that
new words cannot enter the model and that you have to know a priori what words
will model the corpus effectively. A setting where this could be problematic is
when analyzing data where new words in the form of slang or names emerge over 
time. Another limitation is the fact that one has to state
the number of topics the model should use. This is not a big issue, but could
lead to some of the resulting topics to be insignificant.
\begin{center}
\begin{figure}[h]
\centering
\includegraphics[width=0.6\textwidth, natwidth=561,natheight=197]{figs/LDA_standard_model.png}
\caption{Plate notation of the standard LDA model}
\end{figure}
\end{center}

@@

The implementation was run on a computer with an Intel® Xeon® Processor E5-1410
(10MB Cache, 2.8 GHz, 4 cores, 8 threads) and RAM=12 GB.

@@

In figure x we show the 10 topics that consists of the 20 top words with the
highest posterior probability of belonging to the topic. The size of each word
is in proportion to each probability; the bigger the word, the higher the
probability. Each wordcloud describes a topic, which we can find intuitively.
Below are listed the topics that we found corresponding to the words that
describe them:
\begin{itemize}
\item[] Topic 0: Molecular biology
\item[] Topic 1: Materials \& Manufacturing
\item[] Topic 2: Social policy
\item[] Topic 3: Disease detection \& treatment
\item[] Topic 4: Materials science / Physics
\item[] Topic 5: Transport \& Energy
\item[] Topic 6: Business \& Innovation
\item[] Topic 7: Climate change / Environment
\item[] Topic 8: Molecular mechanisms
\item[] Topic 9: Information systems
\item[] Topic 10: International research
\item[] Topic 11: Energy production
\end{itemize}
\begin{center}
\begin{figure}
\begin{tabular}{rcl}
\subfloat[Topic 0]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/EU_topic0_all.png}} &
\subfloat[Topic 1]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/EU_topic1_all.png}} &
\subfloat[Topic 2]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/EU_topic2_all.png}}\\
\subfloat[Topic 3]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/EU_topic3_all.png}} &
\subfloat[Topic 4]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/EU_topic4_all.png}} &
\subfloat[Topic 5]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/EU_topic5_all.png}}\\
\subfloat[Topic 6]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/EU_topic6_all.png}} &
\subfloat[Topic 7]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/EU_topic7_all.png}} &
\subfloat[Topic 8]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/EU_topic8_all.png}}\\
\subfloat[Topic 9]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/EU_topic9_all.png}} &
\subfloat[Topic 10]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/EU_topic10_all.png}} &
\subfloat[Topic 11]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/EU_topic11_all.png}}\\
\end{tabular}
\caption{Wordclouds of topics of the EU dataset (1994-2017)}
\end{figure}
\end{center}
In order to find the topic distribution over time, we need the topic
distribution of each project-document. As we mentioned before, LDA assumes that
each document exhibits multiple topics. After running the model to our dataset,
we get the document-topic distribution and we keep the topic with the highest
probability (the one that most likely belongs to this document). Then, we
calculate the proportion of each topic per FP, using the following equation:
\begin{equation}
topicProportion(k, fp) = \frac{\sum num of docs_{k,fp}}{\sum num of docs_{fp}} * 100
\end{equation}
where $k$ is the number of the topic id, $fp$ is the framework programme,
$fp \in [FP4, FP5, FP6, FP7, H2020]$. The topic proportion of each topic in 
the FP equals to the sum of the number of the documents assigned to this 
topic during the FP, normalized by the number of the documents-projects 
occurring in this FP. In figure X we plot the results of the topic proportion 
over time.

\begin{center}
\begin{figure}
\includegraphics[width=1.0\textwidth, natwidth=707,natheight=483]
{figs/eu-topic-evolution.png}
\caption{Topics over time in EU}
\end{figure}
\end{center}

\subparagraph{Trends in the USA	}

@@

\begin{center}
\begin{figure}
\centering
\begin{tabular}{rcl}
\subfloat[Topic 0]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/usa_topics/USA_topic0_12000_all_docs.png}} &
\subfloat[Topic 1]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/usa_topics/USA_topic1_12000_all_docs.png}} &
\subfloat[Topic 2]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/usa_topics/USA_topic2_12000_all_docs.png}}\\
\subfloat[Topic 3]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/usa_topics/USA_topic3_12000_all_docs.png}} &
\subfloat[Topic 4]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/usa_topics/USA_topic4_12000_all_docs.png}} &
\subfloat[Topic 5]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/usa_topics/USA_topic5_12000_all_docs.png}}\\
\subfloat[Topic 6]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/usa_topics/USA_topic6_12000_all_docs.png}} &
\subfloat[Topic 7]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/usa_topics/USA_topic7_12000_all_docs.png}} &
\subfloat[Topic 8]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/usa_topics/USA_topic8_12000_all_docs.png}}\\
\subfloat[Topic 9]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/usa_topics/USA_topic9_12000_all_docs.png}} &
\end{tabular}
\caption{Wordclouds of topics of the USA dataset (1994-2016)}
\end{figure}
\end{center}

@@

In general, the STEM (science, technology, engineering and math) related
research and education is becoming more and more popular, which is not
surprising, since the National Science Foundation itself estimates that 80/% of
the jobs available during the next decade will require math and science skills.
Even the concept of STEM was first introduced by Judith A. Ramaley, the former
director of the National Science Foundation’s education and human resources
division.

\begin{center}
\begin{figure}
\includegraphics[width=1.0\textwidth, natwidth=707,natheight=483]
{figs/topic-evolution-usa.png}
\caption{Topics over time in USA}
\end{figure}
\end{center}

@@

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth, natwidth=1080,natheight=720]
{figs/heatmaps/heatmapFP4.png}
\caption{Heatmap of the similarity of the two regions during 1994-1997}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth, natwidth=1080,natheight=720]
{figs/heatmaps/heatmapFP5.png}
\caption{Heatmap of the similarity of the two regions during 1998-2001}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth, natwidth=1080,natheight=720]
{figs/heatmaps/heatmapFP6.png}
\caption{Heatmap of the similarity of the two regions during 2002-2006}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth, natwidth=1080,natheight=720]
{figs/heatmaps/heatmapFP7.png}
\caption{Heatmap of the similarity of the two regions during 2007-2013}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth, natwidth=1080,natheight=720]
{figs/heatmaps/heatmapH2020.png}
\caption{Heatmap of the similarity of the two regions during 2014-2017}
\end{figure}

@@

\clearpage
\appendix

\begin{center}
\begin{figure}
\begin{tabular}{rcl}
\subfloat[Quality Management Systems]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/eu_FP4_wordclouds/EUFP4_topic0_7000.png}} &
\subfloat[Agriculture, Fisheries and Food]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/eu_FP4_wordclouds/EUFP4_topic1_7000.png}}&
\subfloat[Molecular biology, Genetics]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/eu_FP4_wordclouds/EUFP4_topic2_7000.png}}\\
\subfloat[EU transport]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/eu_FP4_wordclouds/EUFP4_topic3_7000.png}}&
\subfloat[Digital systems]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/eu_FP4_wordclouds/EUFP4_topic4_7000.png}} &
\subfloat[Materials processing]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/eu_FP4_wordclouds/EUFP4_topic5_7000.png}}\\
\subfloat[Genetics]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/eu_FP4_wordclouds/EUFP4_topic6_7000.png}} &
\subfloat[Policy for environmental sustainability]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/eu_FP4_wordclouds/EUFP4_topic7_7000.png}}&
\subfloat[Energy technologies]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/eu_FP4_wordclouds/EUFP4_topic8_7000.png}}\\
\subfloat[Applications in mobile networks]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/eu_FP4_wordclouds/EUFP4_topic9_7000.png}}&
\end{tabular}
\caption{Wordclouds of topics of the EU FP4 (1994-1998)}
\end{figure}
\end{center}

\begin{center}
\begin{figure}
\begin{tabular}{rcl}
\subfloat[Management software]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/eu_FP5_wordclouds/EUFP5_topic0_7000.png}} &
\subfloat[Hydropower / Energy production]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/eu_FP5_wordclouds/EUFP5_topic1_7000.png}}&
\subfloat[Information technologies]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/eu_FP5_wordclouds/EUFP5_topic2_7000.png}}\\
\subfloat[Quality control systems]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/eu_FP5_wordclouds/EUFP5_topic3_7000.png}}&
\subfloat[Health care / Medical genetics]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/eu_FP5_wordclouds/EUFP5_topic4_7000.png}} &
\subfloat[Environmental sustainability and policies]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/eu_FP5_wordclouds/EUFP5_topic5_7000.png}}\\
\subfloat[Molecular biology (genes)]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/eu_FP5_wordclouds/EUFP5_topic6_7000.png}} &
\subfloat[Optical quantum computing]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/eu_FP5_wordclouds/EUFP5_topic7_7000.png}}&
\subfloat[Education in Nuclear physics]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/eu_FP5_wordclouds/EUFP5_topic8_7000.png}}\\
\subfloat[Education, training]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/eu_FP5_wordclouds/EUFP5_topic9_7000.png}}&
\end{tabular}
\caption{Wordclouds of topics of the EU FP5 (1998-2002)}
\end{figure}
\end{center}

\begin{center}
\begin{figure}
\begin{tabular}{rcl}
\subfloat[International research]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/eu_FP6_wordclouds/EUFP6_topic0_7000.png}} &
\subfloat[Biology]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/eu_FP6_wordclouds/EUFP6_topic1_7000.png}}&
\subfloat[Embedded systems design]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/eu_FP6_wordclouds/EUFP6_topic2_7000.png}}\\
\subfloat[Molecular biology, genes]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/eu_FP6_wordclouds/EUFP6_topic3_7000.png}}&
\subfloat[Evolution / Formation]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/eu_FP6_wordclouds/EUFP6_topic4_7000.png}} &
\subfloat[Sustainable production / Energy]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/eu_FP6_wordclouds/EUFP6_topic5_7000.png}}\\
\subfloat[Economic \& Social policy]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/eu_FP6_wordclouds/EUFP6_topic6_7000.png}} &
\subfloat[Environmental management services]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/eu_FP6_wordclouds/EUFP6_topic7_7000.png}}&
\subfloat[Materials]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/eu_FP6_wordclouds/EUFP6_topic8_7000.png}}\\
\subfloat[Cognitive neuroscience]{\includegraphics[width=0.35\textwidth, natwidth=640,natheight=480]{figs/eu_topics/eu_FP6_wordclouds/EUFP6_topic9_7000.png}}&
\end{tabular}
\caption{Wordclouds of topics of the EU FP6 (2002-2006)}
\end{figure}
\end{center}

>>>>

\setlength{\parskip}{1em}
\usepackage[justification=centering]{caption}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{cite}
\usepackage{algorithm2e}
\usepackage{url}
\usepackage{subfig}
\usepackage{listings}
\usepackage[utf8]{inputenc}
\usepackage{color}

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

\lstdefinestyle{mystyle}{
    backgroundcolor=\color{backcolour},   
    commentstyle=\color{codegreen},
    keywordstyle=\color{magenta},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codepurple},
    basicstyle=\footnotesize,
    breakatwhitespace=false,         
    breaklines=true,                 
    captionpos=b,                    
    keepspaces=true,                 
    numbers=left,                    
    numbersep=5pt,                  
    showspaces=false,                
    showstringspaces=false,
    showtabs=false,                  
    tabsize=2
}

\lstset{style=mystyle}

@@

\begin{enumerate}
\item Pick a document \textbf{w} with probability $p($\textbf{w}$)$
\item For each of the $N$ words in \textbf{w}:
\begin{description}
\item (a) Pick a topic $z$ with probability $p(z|$\textbf{w}$)$
\end{description}
\begin{description}
\item (b) Generate a word $w$ with probability $p(w|z)$
\end{description}
\end{enumerate}
\begin{center}
\begin{figure}[h]
\centering
\includegraphics[width=0.6\textwidth]{figs/PLSI_plate_notation.png}
\caption{Plate notation representing the PLSI model. \textbf{w} is the document 
index, $z$ is the topic drawn for the word $w$ from the topic distribution of 
the document}
\end{figure}
\end{center}
The result from PLSI is that every document is represented as mixing proportions
for the topics, given by $p(z|$\textbf{w}$)$. Even though PLSI is generative for 
the analyzed corpus, it is not generative for new documents, which means is that
there is no clear way of assigning probability to a document that is not part
of the training data.\cite{blei2003latent} Another problem is that the number 
of parameters in the model grows linearly with the number of documents.

@@

A limitation with the standard LDA model is the fix vocabulary. This means that
new words cannot enter the model and that you have to know a priori what words
will model the corpus effectively. A setting where this could be problematic is
when analyzing data where new words in the form of slang or names emerge over 
time. Another limitation is the fact that one has to state
the number of topics the model should use. This is not a big issue, but could
lead to some of the resulting topics to be insignificant.
\begin{center}
\begin{figure}[h]
\centering
\includegraphics[width=0.6\textwidth]{figs/LDA_standard_model.png}
\caption{Plate notation of the standard LDA model}
\end{figure}
\end{center}

@@

The implementation was run on a computer with an Intel® Xeon® Processor E5-1410
(10MB Cache, 2.8 GHz, 4 cores, 8 threads) and RAM=12 GB. Check the Appendix for 
the actual implementation.

@@

In figure x we show the 10 topics that consists of the 20 top words with the
highest posterior probability of belonging to the topic. The size of each word
is in proportion to each probability; the bigger the word, the higher the
probability. Each wordcloud describes a topic, which we can find intuitively.
Below are listed the topics that we found corresponding to the words that
describe them:
\begin{itemize}
\item[] Topic 0: Molecular biology
\item[] Topic 1: Materials \& Manufacturing
\item[] Topic 2: Social policy
\item[] Topic 3: Disease detection \& treatment
\item[] Topic 4: Materials science / Physics
\item[] Topic 5: Transport \& Energy
\item[] Topic 6: Business \& Innovation
\item[] Topic 7: Climate change / Environment
\item[] Topic 8: Molecular mechanisms
\item[] Topic 9: Information systems
\item[] Topic 10: International research
\item[] Topic 11: Energy production
\end{itemize}
\begin{center}
\begin{figure}
\begin{tabular}{rcl}
\subfloat[Topic 0]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/EU_topic0_all.png}} &
\subfloat[Topic 1]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/EU_topic1_all.png}} &
\subfloat[Topic 2]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/EU_topic2_all.png}}\\
\subfloat[Topic 3]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/EU_topic3_all.png}} &
\subfloat[Topic 4]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/EU_topic4_all.png}} &
\subfloat[Topic 5]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/EU_topic5_all.png}}\\
\subfloat[Topic 6]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/EU_topic6_all.png}} &
\subfloat[Topic 7]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/EU_topic7_all.png}} &
\subfloat[Topic 8]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/EU_topic8_all.png}}\\
\subfloat[Topic 9]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/EU_topic9_all.png}} &
\subfloat[Topic 10]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/EU_topic10_all.png}} &
\subfloat[Topic 11]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/EU_topic11_all.png}}\\
\end{tabular}
\caption{Wordclouds of topics of the EU dataset (1994-2017)}
\end{figure}
\end{center}
In order to find the topic distribution over time, we need the topic
distribution of each project-document. As we mentioned before, LDA assumes that
each document exhibits multiple topics. After running the model to our dataset,
we get the document-topic distribution and we keep the topic with the highest
probability (the one that most likely belongs to this document). Then, we
calculate the proportion of each topic per FP, using the following equation:
\begin{equation}
topicProportion(k, fp) = \frac{\sum num of docs_{k,fp}}{\sum num of docs_{fp}} * 100
\end{equation}
where $k$ is the number of the topic id, $fp$ is the framework programme,
$fp \in [FP4, FP5, FP6, FP7, H2020]$. The topic proportion of each topic in 
the FP equals to the sum of the number of the documents assigned to this 
topic during the FP, normalized by the number of the documents-projects 
occurring in this FP. In figure X we plot the results of the topic proportion 
over time.

Based on this plot, topic 10, which refers to International Research, has an
interesting evolution; During the years 1994-1998 is appears as less popular
topic, compared to other topics. However, this state changes drastically in
1998-2002 ($+14\%$) and in 2002-2006 it is at its peak ($+18\% since 1994$).
This period many activities and projects where initialized in order to support
and promote international education and conferences to attract researchers from
all over the world. Strangely, this trend has a significant decrease the last 10
years, compared to other topics ($-12\%$). We already know that in Horizon 2020
there is an agreement between EU and USA for supporting the cooperation of the
two regions.Our hypothesis for the evolution of the International Research is
that the EU successfully managed to create the foundation for supporting
international people and there is no need to invest more on projects that
attempt to promote cooperation between international researchers, because it is
already considered as a common practice nowadays.

Materials and Manufacturing was a hot topic in 1994-1998, but it also appears a
big decrease during the years 1998-2006 and it has a stable trend since then.
Materials Science/Chemistry \& Physics' popularity (topic 4) line has an
increase during 2002-2013 and the last 3 years a small decrease of $2-3\%$.
Topics 2 and 3 (Social policy, Disease detection \& treatment) follow the same
linear trend as Materials science \& Physics. Energy production (topic 11) was
less popular in 1994-2002 and is becoming slightly popular the last 15 years.

Topic 9, which refers to Information systems in general, is a subject that has a
stable trend and it appears as a popular topic the last 22 years compared to
other topics, with a percentage of $12.5\%$ of the projects in EU on average.

Biology and the study of molecular mechanisms have a small decrease of
popularity on average. This may happen because the projects related to medical
fields are supported by another institute, National Institutes of Health. Also,
research in Climate change and Environmental science is another subject that
paradoxically appears as a topic that is gradually losing its popularity over
the last years.

Finally, topic 6, which is related to Business \& Innovation, has a huge
increase the last 3 years. We can assume that the EU is very interested in
supporting the economy by funding projects about businesses and monetary
affairs.
\begin{center}
\begin{figure}
\includegraphics[width=1.0\textwidth]
{figs/eu-topic-evolution.png}
\caption{Topics over time in EU}
\end{figure}
\end{center}
\subparagraph{Trends in the USA	}

@@

\begin{center}
\begin{figure}
\centering
\begin{tabular}{rcl}
\subfloat[Topic 0]{\includegraphics[width=0.35\textwidth]{figs/usa_topics/USA_topic0_12000_all_docs.png}} &
\subfloat[Topic 1]{\includegraphics[width=0.35\textwidth]{figs/usa_topics/USA_topic1_12000_all_docs.png}} &
\subfloat[Topic 2]{\includegraphics[width=0.35\textwidth]{figs/usa_topics/USA_topic2_12000_all_docs.png}}\\
\subfloat[Topic 3]{\includegraphics[width=0.35\textwidth]{figs/usa_topics/USA_topic3_12000_all_docs.png}} &
\subfloat[Topic 4]{\includegraphics[width=0.35\textwidth]{figs/usa_topics/USA_topic4_12000_all_docs.png}} &
\subfloat[Topic 5]{\includegraphics[width=0.35\textwidth]{figs/usa_topics/USA_topic5_12000_all_docs.png}}\\
\subfloat[Topic 6]{\includegraphics[width=0.35\textwidth]{figs/usa_topics/USA_topic6_12000_all_docs.png}} &
\subfloat[Topic 7]{\includegraphics[width=0.35\textwidth]{figs/usa_topics/USA_topic7_12000_all_docs.png}} &
\subfloat[Topic 8]{\includegraphics[width=0.35\textwidth]{figs/usa_topics/USA_topic8_12000_all_docs.png}}\\
\subfloat[Topic 9]{\includegraphics[width=0.35\textwidth]{figs/usa_topics/USA_topic9_12000_all_docs.png}} &
\end{tabular}
\caption{Wordclouds of topics of the USA dataset (1994-2016)}
\end{figure}
\end{center}

@@

In general, the STEM (science, technology, engineering and math) related
research and education is becoming more and more popular, which is not
surprising, since the National Science Foundation itself estimates that 80/% of
the jobs available during the next decade will require math and science skills.
Even the concept of STEM was first introduced by Judith A. Ramaley, the former
director of the National Science Foundation’s education and human resources
division.
\begin{center}
\begin{figure}
\includegraphics[width=1.0\textwidth]
{figs/topic-evolution-usa.png}
\caption{Topics over time in USA}
\end{figure}
\end{center}

@@

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]
{figs/heatmaps/heatmapFP4.png}
\caption{Heatmap of the similarity of the two regions during 1994-1997}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]
{figs/heatmaps/heatmapFP5.png}
\caption{Heatmap of the similarity of the two regions during 1998-2001}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]
{figs/heatmaps/heatmapFP6.png}
\caption{Heatmap of the similarity of the two regions during 2002-2006}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]
{figs/heatmaps/heatmapFP7.png}
\caption{Heatmap of the similarity of the two regions during 2007-2013}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]
{figs/heatmaps/heatmapH2020.png}
\caption{Heatmap of the similarity of the two regions during 2014-2017}
\end{figure}

@@

\clearpage
\section{Appendix}

\textbf{eu iterations.py}

\lstinputlisting[language=Python]{code/eu_iterations.py}

\textbf{usa iterations.py}

\lstinputlisting[language=Python]{code/usa_iterations.py}

\begin{center}
\begin{figure}[h]
\begin{tabular}{rcl}
\subfloat[Quality Management Systems]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP4_wordclouds/EUFP4_topic0_7000.png}} &
\subfloat[Agriculture, Fisheries and Food]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP4_wordclouds/EUFP4_topic1_7000.png}}&
\subfloat[Molecular biology, Genetics]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP4_wordclouds/EUFP4_topic2_7000.png}}\\
\subfloat[EU transport]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP4_wordclouds/EUFP4_topic3_7000.png}}&
\subfloat[Digital systems]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP4_wordclouds/EUFP4_topic4_7000.png}} &
\subfloat[Materials processing]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP4_wordclouds/EUFP4_topic5_7000.png}}\\
\subfloat[Genetics]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP4_wordclouds/EUFP4_topic6_7000.png}} &
\subfloat[Policy for environmental sustainability]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP4_wordclouds/EUFP4_topic7_7000.png}}&
\subfloat[Energy technologies]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP4_wordclouds/EUFP4_topic8_7000.png}}\\
\subfloat[Applications in mobile networks]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP4_wordclouds/EUFP4_topic9_7000.png}}&
\end{tabular}
\caption{Wordclouds of topics of the EU FP4 (1994-1998)}
\end{figure}
\end{center}

\begin{center}
\begin{figure}[h]
\begin{tabular}{rcl}
\subfloat[Management software]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP5_wordclouds/EUFP5_topic0_7000.png}} &
\subfloat[Hydropower / Energy production]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP5_wordclouds/EUFP5_topic1_7000.png}}&
\subfloat[Information technologies]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP5_wordclouds/EUFP5_topic2_7000.png}}\\
\subfloat[Quality control systems]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP5_wordclouds/EUFP5_topic3_7000.png}}&
\subfloat[Health care / Medical genetics]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP5_wordclouds/EUFP5_topic4_7000.png}} &
\subfloat[Environmental sustainability and policies]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP5_wordclouds/EUFP5_topic5_7000.png}}\\
\subfloat[Molecular biology (genes)]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP5_wordclouds/EUFP5_topic6_7000.png}} &
\subfloat[Optical quantum computing]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP5_wordclouds/EUFP5_topic7_7000.png}}&
\subfloat[Education in Nuclear physics]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP5_wordclouds/EUFP5_topic8_7000.png}}\\
\subfloat[Education, training]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP5_wordclouds/EUFP5_topic9_7000.png}}&
\end{tabular}
\caption{Wordclouds of topics of the EU FP5 (1998-2002)}
\end{figure}
\end{center}

\begin{center}
\begin{figure}[h]
\begin{tabular}{rcl}
\subfloat[International research]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP6_wordclouds/EUFP6_topic0_7000.png}} &
\subfloat[Biology]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP6_wordclouds/EUFP6_topic1_7000.png}}&
\subfloat[Embedded systems design]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP6_wordclouds/EUFP6_topic2_7000.png}}\\
\subfloat[Molecular biology, genes]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP6_wordclouds/EUFP6_topic3_7000.png}}&
\subfloat[Evolution / Formation]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP6_wordclouds/EUFP6_topic4_7000.png}} &
\subfloat[Sustainable production / Energy]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP6_wordclouds/EUFP6_topic5_7000.png}}\\
\subfloat[Economic \& Social policy]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP6_wordclouds/EUFP6_topic6_7000.png}} &
\subfloat[Environmental management services]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP6_wordclouds/EUFP6_topic7_7000.png}}&
\subfloat[Materials]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP6_wordclouds/EUFP6_topic8_7000.png}}\\
\subfloat[Cognitive neuroscience]{\includegraphics[width=0.35\textwidth]{figs/eu_topics/eu_FP6_wordclouds/EUFP6_topic9_7000.png}}&
\end{tabular}
\caption{Wordclouds of topics of the EU FP6 (2002-2006)}
\end{figure}
\end{center}
